---
title: "scanorama_integration_simplified"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '/repos/P02567_TBIO-3021_10X_pilot/') # markdown scripts change the working directory to be this script's directory. Change back to head directory of repo
```


Set up the R environment and launch Python
```{r r_env}
`%>%` <- magrittr::`%>%`
getwd()
library(reticulate)
library(Seurat)
Sys.setenv("RETICULATE_PYTHON"="XXXX")

use_condaenv("Scanorama",
             "XXXX",
             require = TRUE)
source("./code/R/Utils/utils_10X_matrix.R")
source("./code/R/Utils/utils_10X_signatures.R")
source("./code/utils/utils_signatures.R")
source("./code/R/Utils/utils_10X_in_out.R")
source("./code/R/Utils/utils_10X_cohort.R")
source("./code/R/Utils/utils_10X_scanorama_integration.R")
source("./code/R/Utils/utils_10X_object.R")

sc <- import("scanpy")
scanorama <- import('scanorama')
```


Set up the python environment
```{python py_env}
import numpy as np
import pandas as pd
import scanpy as sc
import matplotlib.pyplot as plt
import matplotlib
import json as json
from pathlib import Path, PurePath
import scanorama
import scanpy.external as sce

sc.settings.verbosity = 1
```

Prep Integration parameters
```{r}
# Number of anchor features
num_features <- 7000
# Use SCT pearson reisduals?
sct <- TRUE
assay <- "SCT"
slot <- "scale.data" # "data" or "scale.data"
knn <- 60
alpha = 0.1
# Adjust future.globals.maxSize for method requirements
options(future.globals.maxSize = 20000*1024^2)



```



List out the the paths for input and outputs
```{r}
input_path <- c("XXXX")
                 
seurat_path <- c("XXXX")

output_path <- "XXXX"
dir.create(output_path)

output_figs <- "XXXX"
dir.create(output_figs)

# meta_df <- readr::read_delim(paste0(input_path, "full_pdac_meta_data.txt"), delim = "\t")

# cohort_regex <- "PDAC_759110QB_ROI2_s1|PDAC_Pt1_ROI2|PDAC_Pt2_ROI3|PDAC_759110QB_ROI4_s1|PDAC_759110QB_ROI4_s2|PDAC_Pt5_ROI3|PDAC_1255908B_s3|PDAC_1255908B_s4|HumanPanc_S001890_FFPE_C_Apr21|HumanPanc_S001891_FFPE_D_Apr21"

# cohort_regex <- "PDAC_759110QB_ROI2_s1|PDAC_Pt1_ROI2|PDAC_Pt2_ROI3|PDAC_759110QB_ROI4_s1|PDAC_Pt5_ROI3|PDAC_1255908B_s3|HumanPanc_S001890_FFPE_C_Apr21|HumanPanc_S001891_FFPE_D_Apr21"

cohort_regex <- "full_pdac_cohort_seurat_noSCT.RDS"
reject_regex <- NULL #"file_list|v2|FF"
# selected <-meta_df %>% dplyr::filter(Quality_Status == TRUE & Primary == TRUE & grepl("Pt", Sample_Name) & !is.na(Slide))

# cohort_regex <- stringr::str_c(selected$Sample_Name, collapse = "|")


# List of misc fields to keep in meta data
misc_field_list <- c("user.Sample_Name", "user.Group_Name", "user.Tissue", 
                     "user.Date", "user.Area", "user.Project_ID", "user.Primary", 
                     "user.Protocol", "user.Organism", "user.pt.size.factor", 
                     "user.Clustering", "user.Slide", "Pathology.Tissue")
```



Import Seurat objects as a list
```{r, warning=FALSE}
# Find all files following the regex
file_list <-  file_list <- dir(seurat_path, pattern = cohort_regex, full.names = TRUE)
if(is.null(reject_regex) == FALSE){
  file_list <- file_list[!grepl(reject_regex, file_list)]
}

# Ingest a set of RDS Seurat objects288
start_time <- proc.time()
seurat_list <- read_rds_list_simple_my(file_list) #oxford megas = 5 minutes
proc.time() - start_time
```

```{r}
vars_to_regress <- c("mito_percent", "ribo_percent", "user.Sample_Name")
# merged_seurat_data <- seurat_list[[1]]

# Perform SCTransform normalization
merged_seurat_data <- Seurat::SCTransform(merged_seurat_data, assay="Spatial", 
                                         vars.to.regress = vars_to_regress, variable.features.n = num_features,
                                         return.only.var.genes = TRUE, verbose = TRUE, conserve.memory = TRUE)
```

```{r}

```


```{r}
  # Update image names and remove previous clustering
for(name in names(seurat_list)){
  names(seurat_list[[name]]@images) <- name
  seurat_list[[name]]@meta.data <- seurat_list[[name]]@meta.data[which(!grepl("SCT_snn_res.", names(seurat_list[[name]]@meta.data)))]
}
# Move misc parameters to meta data slot
seurat_list <- seurat_misc_to_meta(seurat_list, misc_field_list=misc_field_list)
```


Or if merged seurat object exists:
```{r}
# filename <- paste0(output_path, "oxford_cohort_merge.RDS")
# filename <- paste0("XXXX")
# data_seurat_merged <- readRDS(filename)
# seurat_list <- SplitObject(data_seurat_merged, "Batch")
```

```{r}
seurat_list <- Seurat::SplitObject(merged_seurat_data, split.by = "user.Sample_Name")
```



Wrangle data for integration
```{r}
exprs <- prep_scanorama_integration_my(seurat_list, num_features = num_features, sct = sct,
                                       slot = slot, assay = assay)

genes <- lapply(exprs, get_gene_names_my)

```


These lists get changed to dictionaries in python. Change these to lists and double check the shape
```{python}
python_exprs = r.exprs
# python_exprs
data = python_exprs.values()
new_list = list(data)
# new_list = [mat.tocsr() for mat in new_list] #this is if the data is a sparse matrix instead of dense, i.e. slot = "data"
new_list[0].shape
# 
gene_data = r.genes.values()
gene_list = list(gene_data)
len(gene_list[0])
```


double check that the data is  alist of numpy.ndarray or scipy csr matrices
```{python}
type(new_list[1])
```


Now integrate the data using scanorama
```{python}
import time
start_time = time.time()
integrated_data = scanorama.integrate(new_list, gene_list, verbose = True, dimred = 50, knn=int(r.knn), alpha = 0.1)
print("--- %s seconds ---" % (time.time() - start_time))
```


Parse out the scanorama embeddeing matrices and concatenate.
```{python}
X_scanorama = integrated_data[0]
scanorama_gene_list = integrated_data[1]

all_s = np.concatenate(X_scanorama)
print(all_s.shape)
```



Keep track of sample IDs in scanorama matrix. (might be able to get rid of this?)
```{r}
get_cell_num_scanorama_list <- function(x_scanorama_list){
  item_dims <- dim(x_scanorama_list)
  return(item_dims[1])
}

nums <- lapply(py$X_scanorama, get_cell_num_scanorama_list) %>% unlist

temp_list <- list()
sample_names <- names(exprs)
for(i in c(1:length(nums))){
  this_sample <- sample_names[i] %>% stringr::str_remove("_all")
  temp_list[i] <- rep(this_sample, nums[i]) %>% list
}
sample_name_scanorama <- unlist(temp_list)
```


Merge the seurat objects into one.
```{r}
data_seurat_merged <- merge(x = seurat_list[[1]], y = seurat_list[2:length(seurat_list)], project = "subset_pathology_cohort", merge.data = TRUE)
# data_seurat_merged <- seurat_merge_my(seurat_list, keep_SCT = TRUE)
```


```{r}
data_seurat_merged@meta.data$Library_ID = sample_name_scanorama
```

```{r}
# embedding <- py$all_s
# list_spot_names <- lapply(seurat_list, colnames) %>% unlist
# rownames(embedding) <- list_spot_names
# scanorama_obj <- Seurat::CreateDimReducObject(embedding, key = "scanorama_", assay ="SCT")
```

```{r}
# scan_umap <- uwot::umap(embedding, metric = "cosine")
```


Add scanorama embedding to merged seurat object
```{r}
embedding <- py$all_s
rownames(embedding) <- rownames(merged_seurat_data@meta.data)

merged_seurat_data[["scanorama"]] <- Seurat::CreateDimReducObject(embedding, key = "scanorama_", assay = "SCT")
```


```{r}
saveRDS(embedding, "XXXX")

embedding <- readRDS("XXXX")
```



```{r}
# saveRDS(data_seurat_merged, file = paste0(output_path, "oxford_cohort_merge.RDS"), compress = FALSE)
```



Run UMAP on scanorama embedding
```{r}
# data_seurat_merged <- Seurat::FindNeighbors(data_seurat_merged, reduction = "scanorama")
merged_seurat_data <- Seurat::RunUMAP(merged_seurat_data, verbose = TRUE, reduction = "scanorama", dims = 1:50)
```

Wrangle metadata
```{r}
block_ids <- merged_seurat_data@meta.data$user.Sample_Name %>% stringr::str_remove("_ROI[1234]") %>% stringr::str_remove("_s[12345678]")
merged_seurat_data@meta.data$Blocks <- block_ids
```


Plot UMAP grouped by sample
```{r, fig.width = 8}
# colors <- RColorBrewer::brewer.pal(10, name = "Set2")
Seurat::DimPlot(merged_seurat_data, reduction = "umap", group.by = "Blocks", shuffle = TRUE)
ggplot2::ggsave(paste0(output_figs, "_integrated_UMAP.png"), width = 15)
```

```{r, fig.width= 20}
Seurat::DimPlot(merged_seurat_data, reduction = "umap", group.by = "Blocks", shuffle = FALSE, raster = FALSE)+ ggplot2::theme(aspect.ratio = 1)
   ggplot2::ggsave(paste0(output_figs, "_integrated_UMAP.png"), width = 20)
```


```{r, fig.width=16}
# Seurat::DimPlot(data_seurat_merged, reduction = "umap",split.by = "Blocks", group.by = "Pathology.Group", ncol = 4)
patchwork::wrap_plots(
  (Seurat::FeaturePlot(merged_seurat_data, reduction = "umap",
                    split.by = "Blocks",
                    features = "Pathology.Epithelium.percent", combine = FALSE)), ncol = 6, guides = "collect")
```


```{r}
Seurat::FeaturePlot(merged_seurat_data, reduction = "umap",
                    features = "Pathology.Epithelium.percent") + ggplot2::theme(aspect.ratio = 1)
Seurat::FeaturePlot(merged_seurat_data, reduction = "umap",
                    features = "Pathology.Epithelium") + ggplot2::theme(aspect.ratio = 1)
```

```{r}
Seurat::FeaturePlot(merged_seurat_data, reduction = "umap",
                    features = "Pathology.Exocrine.percent") + ggplot2::theme(aspect.ratio = 1)
```

```{r}
Seurat::FeaturePlot(merged_seurat_data, reduction = "umap",
                    features = "Pathology.Benign Glands.percent") + ggplot2::theme(aspect.ratio = 1)
```

```{r}
Seurat::FeaturePlot(merged_seurat_data, reduction = "umap",
                    features = "Pathology.Stroma.percent") + ggplot2::theme(aspect.ratio = 1)
```

```{r}
Seurat::FeaturePlot(merged_seurat_data, reduction = "umap",
                    features = "Pathology.Stroma") + ggplot2::theme(aspect.ratio = 1)
```


```{r}
Seurat::DimPlot(merged_seurat_data, reduction = "umap",group.by = "Pathology.Group", split.by = "Pathology.Group", ncol = 4) + ggplot2::theme(aspect.ratio = 1)
```



```{r, fig.width=15}
Seurat::DimPlot(data_seurat_merged, reduction = "umap", group.by = "Blocks")
ggplot2::ggsave(paste0(output_figs, "_integrated_UMAP_BlockID.png"), width = 15)
```


UMAP faceted by sample.
```{r, fig.width=18}
# colors <- RColorBrewer::brewer.pal(8, name = "Set2")
Seurat::DimPlot(data_seurat_merged, reduction = "umap", split.by = "Library_ID", group.by = "Blocks", ncol = 8, raster = FALSE)+ ggplot2::theme(aspect.ratio = 1)
ggplot2::ggsave(paste0(output_figs, "integrated_UMAP_facet.png"))
```

Add common signatures to seurat object
```{r}
sig_select <- c("PDAC.collisson.classical","PDAC.moffitt.basal", "PDAC.P19.Fibroblast", "PDAC.U.Exocrine.Acinar", "PDAC.U.Endocrine.Beta", "PDAC.moffitt.normalstroma", "PDAC.U.Immune.Tcell.CD8", "PDAC.U.Immune.Bcell", "PDAC.U.Immune.Macrophage")
# Path to a signature file
sig_path <- "/repos/P02567_TBIO-3021_10X_pilot/data/import/Signatures/signatures_220210.txt"

sct_threshold <- 0.25
spot_threshold <- 5
assay = 'SCT'
slot = 'data'
# Find genes abundant in this sample

# gene_list <- seurat_select_abundant_genes_my(data_seurat_merged, sct_threshold=sct_threshold, spot_threshold=spot_threshold, assay=assay, slot=slot)
gene_list <- rownames(merged_seurat_data)
# Load signatures and filter them down to only well represented genes
signature_list <- read_filter_signatures_my(sig_path, gene_list, sig_names=sig_select,
                                          sig_length_min=1, sig_length_max=1000, ratio_threshold=0)
# Add signature scores to a seurat object
merged_seurat_data <- add_signature_scores_my(merged_seurat_data, signature_list, prefix="sig.", assay=assay)

# Seurat renames column names, this step finds new signature names
sig_names_upd <- colnames(merged_seurat_data@meta.data[grep("sig.", colnames(merged_seurat_data@meta.data))])

```

Plot signatures on UMAP
```{r, fig.width=5, fig.height=5}
for(i in c(1:9)){
   Seurat::FeaturePlot(merged_seurat_data, features = sig_names_upd[i], min.cutoff = 0, max.cutoff = 1.5) %>% print()
    ggplot2::ggsave(paste0(output_figs, sig_names_upd[i], "_integrated.png"), width = 5, height = 5)
}
```


Perform leiden clustering embedding.
```{r}
start_time <- proc.time()
merged_seurat_data <- Seurat::FindNeighbors(merged_seurat_data, reduction = "scanorama", dims = 1:50, force.recalc = TRUE, verbose = TRUE)
proc.time() - start_time
start_time <- proc.time()
merged_seurat_data <- Seurat::FindClusters(merged_seurat_data, resolution = 0.5, algorithm = 1, graph.name = "SCT_snn", verbose = TRUE) #48 minutes oxford megas (32 samples)
proc.time() - start_time
```

```{r}
table(merged_seurat_data@meta.data$seurat_clusters, merged_seurat_data@meta.data$Pathology.Group)
table(merged_seurat_data@meta.data$seurat_clusters, merged_seurat_data@meta.data$Blocks)
```

```{r}
cluster_by_sample <- table(merged_seurat_data@meta.data$seurat_clusters, merged_seurat_data@meta.data$user.Sample_Name)
cluster_by_sample <- tibble::as_tibble(cluster_by_sample, .name_repair = "unique")
colnames(cluster_by_sample) <- c("Cluster", "Sample", "Count")
```

```{r}
spot.theme <- list(
  theme_classic(),
  # theme(axis.ticks.x=element_blank(), axis.text.x=element_text(size = 19, angle = 90, hjust = 0)),
  # theme(axis.ticks.y=element_blank(), axis.text.y=element_text(size = 19)),
  theme(axis.line=element_blank()),
  # theme(text = element_text(size = 22)),
  # theme(legend.position = "none"),
  # theme(plot.margin = unit(c(10,10,10,10), "mm")),
  scale_size_continuous(range = c(-0.3, 15)),
  scale_x_discrete(position = "top"))
```



```{r, fig.width=20}
library(ggplot2)
ggplot(cluster_by_sample, aes(x = Cluster, y = Sample))+
  geom_point(aes(size = Count))+
  spot.theme
```


```{r, fig.width=15}
Seurat::DimPlot(merged_seurat_data, group.by = "seurat_clusters", shuffle = TRUE, raster = FALSE,
                reduction = "umap") + ggplot2::theme(aspect.ratio = 1)
ggplot2::ggsave(paste0(output_figs,  "louvain_0.5_integrated_clusters.png"), width = 15, height = 15)
```


```{r, fig.height=20}
cluster_1 <- subset(data_seurat_merged, idents = 1)
Seurat::DimPlot(cluster_1,
                split.by = "Library_ID",
                shuffle = TRUE, raster = TRUE,
                reduction = "umap", ncol = 8) + ggplot2::theme(aspect.ratio = 1)
```

```{r}
start_time <- proc.time()
saveRDS(data_seurat_merged, file = paste0(output_path, "leidein_0.5_oxford_megas_integrated.RDS"), compress = FALSE)
proc.time() - start_time
```


```{r}
table(data_seurat_merged$Library_ID, data_seurat_merged$seurat_clusters)
```



```{r, fig.width=25}
patchwork::wrap_plots(Seurat::SpatialDimPlot(data_seurat_merged, cells.highlight = CellsByIdentities(object = data_seurat_merged, idents = c(4)), combine = FALSE), ncol = 6)
```



```{r, fig.width=20}
cancer_2 <- subset(data_seurat_merged, idents = 5)
Seurat::SpatialDimPlot(cancer_2, crop = TRUE, combine = FALSE) %>% patchwork::wrap_plots(ncol = 6)
Seurat::SpatialFeaturePlot(cancer_2, features = "sig.PDAC.collisson.classical", crop = FALSE)
# Seurat::SpatialDimPlot(data_seurat_merged, crop = TRUE, label = FALSE, images = "slice1.7", pt.size.factor = 1.5) + ggplot2::theme(legend.key.height = ggplot2::unit(20, 'pt'))
```


