---
title: "scanorama integration"
output: html_document
author: "Andy Kavran"
date: "12 April 2022"
---

Set the R environment

Before running this script it seems necessary to run these commands in terminal:

$ cd /XXXX
$ miniconda3/bin/conda init bash
$ bash
$ conda activate Scanorama

it fails at use_condaenv() if I don't do this, but only sometimes....
```{r r_env}
`%>%` <- magrittr::`%>%`
library(reticulate)
Sys.setenv("RETICULATE_PYTHON"="XXXX")

use_condaenv("Scanorama",
             "XXXX",
             require = TRUE)

source("/repos/P02567_TBIO-3021_10X_pilot/code/utils/utils_in_out.R")
source("/repos/P02567_TBIO-3021_10X_pilot/code/R/Utils/utils_10X_vis.R")

sc <- import("scanpy")
```


Set up the python environment
```{python py_env}
import numpy as np
import pandas as pd
import scanpy as sc
import matplotlib.pyplot as plt
import matplotlib
import json as json
from pathlib import Path, PurePath
import scanorama
import scanpy.external as sce

sc.settings.verbosity = 3
```


```{r parameters}
# cohort_regex <- "PDAC_759110QB_ROI2_s1|PDAC_Pt1_ROI2|PDAC_Pt2_ROI3|PDAC_759110QB_ROI4_s1|PDAC_759110QB_ROI4_s2|PDAC_Pt5_ROI3|PDAC_1255908B_s3|PDAC_1255908B_s4|HumanPanc_S001890_FFPE_C_Apr21|HumanPanc_S001891_FFPE_D_Apr21"
cohort_regex <- "Pt|E27038"
input_path <-"XXXX"
output_figs <- "XXXX"
dir.create(output_figs, showWarnings = FALSE)

list_files <- dir(input_path, pattern = paste0(cohort_regex, ".*.h5ad"), full.names = TRUE)
list_files
```

Ingest first dataset, then append each dataset to it
```{python ingest h5 datasets}
import scanpy as sc
adata = []
# X_list = []
for i, file in enumerate(r.list_files):
    temp_data = sc.read_h5ad(file)
    adata.append(temp_data)
    # X_list.append(temp_data.X)
```

```{python}
adata
```


```{python concatenate scanpy objects}
library_names = []
for i in range(len(adata)):
    this_name = adata[i].uns['misc']['user.Sample_Name']
    library_names.append(this_name)
    

adata_cat = adata[0].copy()

adata_cat = adata_cat.concatenate(
    adata[1:len(adata)],
    batch_key="library_id",
    uns_merge="unique",
    batch_categories=library_names)

```

```{python}
adata_cat
```


```{python}
sc.pp.highly_variable_genes(adata_cat, flavor="seurat", n_top_genes=2000, inplace=True, batch_key="library_id")

split = []
batch_categories = adata_cat.obs["library_id"].unique()
for i in batch_categories:
    split.append(adata_cat[adata_cat.obs["library_id"] == i].copy())
# 
for i in range(len(split)):
    sc.pp.scale(split[i])

```

```{python}
# adatas_cor = scanorama.correct_scanpy(split, return_dimred=True)
scanorama.integrate_scanpy(split)
```


```{python}
# sc.tl.pca(split, svd_solver='arpack')
split
```


```{python}
adata_cat
```

Create Seurat Object
```{r}
# Get the expression matrix
exprs <- py$adata_cat$X$todense()
exprs <- t(exprs)
colnames(exprs) <- py$adata_cat$obs_names$to_list()
rownames(exprs) <- py$adata_cat$var_names$to_list()
# Create the Seurat object
seurat <- Seurat::CreateSeuratObject(exprs)
# Set the expression assay
seurat <- Seurat::SetAssayData(seurat, "data", exprs)
# Add observation metadata
seurat <- Seurat::AddMetaData(seurat, py$adata_cat$obs)
# Add embedding
# embedding <- py$adata_cat$obsm["X_pca"]
# rownames(embedding) <- py$adata_cat$obs_names$to_list()
# seurat[["pca"]] <- Seurat::CreateDimReducObject(embedding, key = "pca_")
```


```{python}
# Get all the integrated matrices.
scanorama_int = [ad.obsm['X_scanorama'] for ad in split]

# make into one matrix.
all_s = np.concatenate(scanorama_int)
print(all_s.shape)

# add to the AnnData object
adata_cat.obsm["Scanorama"] = all_s
```


```{r}
# Add embedding
embedding <- py$adata_cat$obsm["Scanorama"]
rownames(embedding) <- py$adata_cat$obs_names$to_list()
# rownames(embedding) <- py$spot_names
seurat[["scanorama"]] <- Seurat::CreateDimReducObject(embedding, key = "scanorama_")

```


```{r}
seurat <- Seurat::FindNeighbors(seurat, reduction = "scanorama", dims = 1:50)
seurat <- Seurat::RunUMAP(seurat, dims = 1:50, verbose = FALSE, reduction = "scanorama")
```

```{r, fig.height=8, fig.width=10}
# Seurat::DimPlot(seurat, reduction = "umap", group.by = "library_id", cols = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e','#e6ab02','#a6761d','#666666'))
# p1 <- Seurat::DimPlot(seurat, reduction = "umap", group.by = "library_id", cols = c('#8dd3c7','#ffffb3','#bebada','#fb8072','#80b1d3','#fdb462','#b3de69','#fccde5','#d9d9d9','#bc80bd'))
p1 <- Seurat::DimPlot(seurat, reduction = "umap", group.by = "library_id")
fig_path <- paste0(output_figs, "/hvg_scaled_scanorama_UMAP")
write_plot2file_my(p1,filename = fig_path, width = 10, height = 8)
```

```{r}
p2 <- Seurat::DimPlot(seurat, reduction = "umap",split.by = "library_id",group.by = "library_id", ncol = 5)+ theme(legend.position = "none")
fig_path <- paste0(output_figs, "/hvg_scaled_scanorama_UMAP_facet")
write_plot2file_my(p2,filename = fig_path, width = 20, height = 8)

```

```{r}
Seurat::DimPlot(seurat, reduction = "umap",split.by = "library_id",group.by = "library_id", ncol = 6)+ theme(legend.position = "none")
```

```{r}
p4 <- Seurat::FeaturePlot(seurat, features = c("DES", "COL1A1", "GATA6","PNLIP"))
fig_file <- paste0(output_figs, "/feature_UMAP")
# write_plot2file_my(p4, filename = fig_file)
```

```{r}
sig_select <- c("PDAC.collisson.classical","PDAC.moffitt.basal", "PDAC.P19.Fibroblast", "PDAC.U.Exocrine.Acinar", "PDAC.U.Endocrine.Beta",
                "PDAC.U.Immune.Tcell.CD8", "PDAC.U.Immune.Bcell", "PDAC.U.Immune.Macrophage", "PDAC.moffitt.activatedstroma", "PDAC.moffitt.normalstroma")
# Path to a signature file
sig_path <- "/repos/P02567_TBIO-3021_10X_pilot/data/import/Signatures/signatures_220210.txt"

sct_threshold <- 0.25
spot_threshold <- 5
assay = 'RNA'
slot = 'data'
# Find genes abundant in this sample

gene_list <- seurat_select_abundant_genes_my(seurat, sct_threshold=sct_threshold, spot_threshold=spot_threshold,
                                               assay=assay, slot=slot)
# Load signatures and filter them down to only well represented genes
signature_list <- read_filter_signatures_my(sig_path, gene_list, sig_names=sig_select,
                                          sig_length_min=1, sig_length_max=1000, ratio_threshold=0)
# Add signature scores to a seurat object
seurat <- add_signature_scores_my(seurat, signature_list, prefix="sig.", assay=assay)

# Seurat renames column names, this step finds new signature names
sig_names_upd <- colnames(seurat@meta.data[grep("sig.", colnames(seurat@meta.data))])

# Set a min and a max value to a Seurat object for a spatial plot
# seurat <- set_spatial_min_max_my(seurat, sig_names_upd)
# sig_names_upd <- colnames(seurat@meta.data[grep("sig.", colnames(seurat@meta.data))])
# sig_num <- sig_names_upd %>% length
```

```{r, fig.width= 7.3, fig.height=6}
# output_figs <- 'XXXX'
for(i in c(1:9)){
   Seurat::FeaturePlot(seurat, features = sig_names_upd[i], min.cutoff = 0, max.cutoff = 1)
    ggsave(paste0(output_figs,"/", sig_names_upd[i], "_integrated.png"))
}
Seurat::FeaturePlot(seurat, features = sig_names_upd[10], min.cutoff = 0, max.cutoff = 2.5)
# Seurat::FeaturePlot(seurat, features = sig_names_upd[1], min.cutoff = 0, max.cutoff = 1)
# ggsave(paste0(sig_names_upd[1], "_integrated.png"))
# Seurat::FeaturePlot(seurat, features = sig_names_upd[2], min.cutoff = 0, max.cutoff = 1)
# ggsave(paste0(sig_names_upd[2], "_integrated.png"))
# Seurat::FeaturePlot(seurat, features = sig_names_upd[3], min.cutoff = 0, max.cutoff = 1)
# ggsave(paste0(sig_names_upd[3], "_integrated.png"))
# Seurat::FeaturePlot(seurat, features = sig_names_upd[4], min.cutoff = 0, max.cutoff = 1)
# ggsave(paste0(sig_names_upd[4], "_integrated.png"))
# Seurat::FeaturePlot(seurat, features = sig_names_upd[5], min.cutoff = 0, max.cutoff = 1)
# ggsave(paste0(sig_names_upd[5], "_integrated.png"))
# Seurat::FeaturePlot(seurat, features = sig_names_upd[6], min.cutoff = 0, max.cutoff = 1)
# ggsave("classical_integrated.png")
# Seurat::FeaturePlot(seurat, features = sig_names_upd[7], min.cutoff = 0, max.cutoff = 1)
# ggsave("classical_integrated.png")
# Seurat::FeaturePlot(seurat, features = sig_names_upd[8], min.cutoff = 0, max.cutoff = 1)
# ggsave("classical_integrated.png")
# Seurat::FeaturePlot(seurat, features = sig_names_upd[9], min.cutoff = 0, max.cutoff = 1)
# ggsave("classical_integrated.png")
# Seurat::FeaturePlot(seurat, features = sig_names_upd[19], min.cutoff = 0, max.cutoff = 1)
# ggsave("classical_integrated.png")
```


```{r}
seurat <- Seurat::FindClusters(seurat, resolution =0.3)
```

```{r}
p3 <- Seurat::DimPlot(seurat, reduction = "umap", group.by ="RNA_snn_res.0.3")
fig_path <- paste0(output_figs, "/scanorama_clusters_UMAP")
write_plot2file_my(p3,filename = fig_path, width = 8, height = 8)
```

```{r}
table(seurat@meta.data$RNA_snn_res.0.3, seurat@meta.data$library_id) %>% image
```


```{python}
# adata_cat.raw = adata_cat
# adata_cat = adata_cat[:,adata_cat.var.highly_variable_nbatches > 1]
# adata_cat
```



```{python}
# adatas = {}
# for batch in library_names:
#     adatas[batch] = adata_cat[adata_cat.obs['library_id'] == batch,]
# 
# adatas 
```


```{python}
# #convert to list of AnnData objects
# adatas = list(adatas.values())
# 
# # run scanorama.integrate
# scanorama.integrate_scanpy(adatas, dimred = 50)
```

```{python}
# # Get all the integrated matrices.
# scanorama_int = [ad.obsm['X_scanorama'] for ad in adatas_cor]
# 
# # make into one matrix.
# all_s = np.concatenate(scanorama_int)
# print(all_s.shape)
# 
# # add to the AnnData object
# adata_cat.obsm["Scanorama"] = all_s
```


```{python}
# spot_names = []
# for i in range(len(adatas_cor)):
#     spot_names.append(adatas_cor[i].obs_names.to_list())
# spot_names
```


```{r}
# # Add embedding
# embedding <- py$adata_cat$obsm["Scanorama"]
# rownames(embedding) <- py$adata_cat$obs_names$to_list()
# # rownames(embedding) <- py$spot_names
# seurat[["scanorama"]] <- Seurat::CreateDimReducObject(embedding, key = "scanorama_")
# # scanorama_mat <- Seurat::CreateDimReducObject(embedding, key = "scanorama_")
```


```{r}
# Seurat::DimPlot(seurat, reduction = "scanorama", group.by = "library_id")
```





```{r}
py$all_s %>% dim

plot(py$all_s[,1], py$all_s[,2])
```


```{r}
Cstack_info()
```


possibly an environment limit. C-stack usage

```{python}
sc.pp.neighbors(adata_cat, use_rep="Scanorama")
sc.tl.umap(adata_cat)
```

```{python}
sc.pl.umap(adata_cat, color = 'library_id')
```


```{python}
sc.pl.umap(adata_cat, color = ["TSPAN8", "COL1A1", "TMSB4X", 'library_id'])
```


```{python}
sc.tl.leiden(adata_cat, resolution = .7, key_added="clusters")
```

```{python}
sc.pl.umap(adata_cat, color = ['clusters'])
```


```{python}
np.unique(adata_cat.obs["library_id"])
adata_cat
```


```{python}
fig, axs = plt.subplots(1, 2, figsize=(15, 10))

for i, library in enumerate(np.unique(adata_cat.obs["library_id"])):
    ad = adata_cat[adata_cat.obs.library_id == library, :].copy()
    sc.pl.spatial(
        ad,
        img_key="hires",
        library_id=library,
        color="clusters",
        size=1.5,
        legend_loc=None,
        show=False,
        ax=axs[i],
    )

plt.tight_layout()
```

```{python}
# for i, library in enumerate(np.unique(adata_cat.obs["library_id"])):
lib = adata_cat.obs.library_id.unique()

ad = adata_cat[adata_cat.obs.library_id == lib[0], :].copy()
sc.pl.spatial(
    ad,
    img_key="hires",
    library_id=lib[0],
    color="clusters",
    size=1.5,
    legend_loc=None,
    show=True)
```


```{python}
# adata_cat
sce.pp.scanorama_integrate(adata_cat, 'library_id', )
```

```{python}
sc.pp.neighbors(adata_cat, use_rep="X_scanorama")
sc.tl.umap(adata_cat)
sc.tl.leiden(adata_cat, key_added="clusters")
```


```{python}
sc.pp.calculate_qc_metrics(adata, percent_top=None, log1p=False, inplace=True)
```

```{python}
sc.pl.violin(adata, ['n_genes_by_counts',  'total_counts'], groupby='library_id', rotation = 45)
```

```{python}
sc.pl.highest_expr_genes(adata, n_top=20)
```

```{python}
# save the counts to a separate object for later, we need the normalized counts in raw for DEG dete
counts_adata = adata.copy()

sc.pp.normalize_total(adata, inplace=True)
sc.pp.log1p(adata)
# take 1500 variable genes per batch and then use the union of them.
sc.pp.highly_variable_genes(adata, flavor="seurat", n_top_genes=1500, inplace=True, batch_key="library_id")

# subset for variable genes
adata.raw = adata
adata = adata[:,adata.var.highly_variable_nbatches > 0]

# scale data
# sc.pp.scale(adata)
```

```{python}
sc.pp.neighbors(adata)
sc.tl.umap(adata)
sc.tl.leiden(adata, key_added="clusters")
```




