---
title: "Seurat to h5ad scanpy objects"
output: html_notebook
author: "Andy Kavran"
date: "7 April 2022"
---

Set the R environment

Before running this script it seems necessary to run these commands in terminal:

$ cd /XXXX
$ miniconda/bin/conda init bash
$ bash
$ conda activate Scanorama

if not it fails at use_condaenv()
```{r r_env}
`%>%` <- magrittr::`%>%`
library(reticulate)
Sys.setenv("RETICULATE_PYTHON"="XXXX")

use_condaenv("Scanorama",
             "XXXX",
             require = TRUE)
# if(!"SeuratDisk"%in% rownames(installed.packages())){
#   devtools::install_git("https://github.com/mojaveazure/seurat-disk.git", dependencies=FALSE)
# }
source("/repos/P02567_TBIO-3021_10X_pilot/code/utils/utils_in_out.R")

sc <- import("scanpy")
```


Set up the python environmnet
```{python py_env}
import numpy as np
import pandas as pd
import scanpy as sc
import matplotlib.pyplot as plt
import matplotlib
import json as json
from pathlib import Path, PurePath

sc.settings.verbosity = 3
```


Parameters for r
```{r r_param}
input_path <- "XXXX"
output_path <-"XXXX"

cohort_regex <- "HumanPanc_S001890_FFPE_C_Apr21|HumanPanc_S001891_FFPE_D_Apr21"

# dir.create(output_path, showWarnings = FALSE)
```


Parameters for python
```{python py_param}
input_path = '/XXXX'
```

Define python functions to create scanpy object from a seurat object
```{python def image function}
def add_visium_image_data(adata, library_id, path):
  # print("testing")
  from matplotlib.image import imread
  from pathlib import Path, PurePath
  from h5py import File
  
  with File(path / "filtered_feature_bc_matrix.h5", mode = "r") as f:
    attrs = dict(f.attrs)

  files = dict(
            tissue_positions_file=path / 'spatial/tissue_positions_list.csv',
            scalefactors_json_file=path / 'spatial/scalefactors_json.json',
            hires_image=path / 'spatial/tissue_hires_image.png',
            lowres_image=path / 'spatial/tissue_lowres_image.png',
        )
  adata.uns['spatial'] = dict()
  adata.uns['spatial'][library_id] = dict()
  adata.uns['spatial'][library_id]['images'] = dict()
  for f in files.values():
    if not f.exists():
        if any(x in str(f) for x in ["hires_image", "lowres_image"]):
            logg.warning(
                f"You seem to be missing an image file.\n"
                f"Could not find '{f}'."
            )
        else:
            raise OSError(f"Could not find '{f}'")
  for res in ['hires', 'lowres']:
    try:
        adata.uns["spatial"][library_id]['images'][res] = imread(
            str(files[f'{res}_image'])
        )
    except Exception:
      raise OSError(f"Could not find '{res}_image'")

  adata.uns["spatial"][library_id]['scalefactors'] = json.loads(
      files['scalefactors_json_file'].read_bytes()
  )
  adata.uns["spatial"][library_id]["metadata"] = {
    k: (str(attrs[k], "utf-8") if isinstance(attrs[k], bytes) else attrs[k])
    for k in ("chemistry_description", "software_version")
    if k in attrs
  }
  positions = pd.read_csv(files['tissue_positions_file'], header=None)
  positions.columns = [
    'barcode',
    'in_tissue',
    'array_row',
    'array_col',
    'pxl_col_in_fullres',
    'pxl_row_in_fullres',
    ]
  positions.index = positions['barcode']

  adata.obs = adata.obs.join(positions, how="left")
  adata.obsm['spatial'] = adata.obs[
  ['pxl_row_in_fullres', 'pxl_col_in_fullres']
  ].to_numpy()
  adata.obs.drop(
    columns=['barcode', 'pxl_row_in_fullres', 'pxl_col_in_fullres'],
    inplace=True,
    )
    
  return adata
  
def scanpy_from_seurat(exprs, meta, feature_meta, misc_meta):
  adata = sc.AnnData(X = exprs, obs = meta, var = feature_meta)
  # adata.obsm['umap'] = embedding_umap
  # adata.obsm['pca'] = embedding_pca
  adata.uns['misc'] = misc_meta
  path = Path(adata.uns["misc"]["user.FullPath"])
  library_id = adata.uns["misc"]["user.Sample_Name"]
  adata = add_visium_image_data(adata, library_id, path)
  
  return adata
```

In R, loop through annotated seurat objects. Ingest seurat objects, then convert and save to h5ad objects.
```{r ingest}
list_files <- dir(input_path, pattern=paste0(cohort_regex, ".*.rds"), full.names=TRUE)

for(file in list_files){
  seurat_data <- readRDS(file)
  this_sample_name <- seurat_data@misc$user.Sample_Name
  exprs <- Seurat::GetAssayData(seurat_data) %>% Matrix::t() # transpose for python
  meta <- seurat_data[[]]
  feature_meta <- Seurat::GetAssay(seurat_data)[[]]
  # embedding_umap <- Seurat::Embeddings(seurat_data, "umap")
  # embedding_pca <- Seurat::Embeddings(seurat_data, "pca")
  misc_meta <- seurat_data@misc
  
  adata <- py$scanpy_from_seurat(exprs, meta, feature_meta, misc_meta)
  output_file <- paste0(output_path, this_sample_name, ".h5ad")
  adata$write(output_file)
  cat("Filed saved:", output_file, "\n")
}

```

Test if a saved filed opens and produces the expected plots.
```{python }
read_adata = sc.read_h5ad("XXXX")
```


```{python}
sc.pl.violin(read_adata, ['nFeature_SCT'], jitter=0.4, rotation= 45)
```


```{python}
sc.pl.violin(read_adata, ['nCount_SCT'], jitter=0.4, rotation= 45)
```
```{python}
sc.pp.highly_variable_genes(read_adata, flavor="seurat", n_top_genes=1500, inplace=True)

```

```{python}
sc.pp.neighbors(read_adata)
```

```{python}
# sc.tl.umap(read_adata)
sc.tl.leiden(read_adata, key_added="clusters",resolution = .4)
```

```{python}
sc.pl.umap(read_adata, color=["clusters"], palette=sc.pl.palettes.default_20)
```


```{python}
sc.pl.spatial(read_adata, img_key="hires", color=["clusters"], size = 1.5)
```

```{python}
sc.pl.umap(read_adata, color=["nFeature_SCT", "nCount_SCT", "SCT_snn_res.0.4"], wspace=0.4)
```

```{python}
sc.pl.umap(read_adata, color=["KRAS"], wspace=0.4)
```

